Sentiment Analysis of Twitter in Relation to Fossil Fuel Stock Prices

### 2023-04-25 manuscript/manuscript.tex
changed

\begin{abstract}
As a publicly traded company, the Chevron Corporation has the 11th highest market capitalization in the world with 361.57 billion dollars in traded stocks. There exist many social factors that contribute to this stock value but this paper aims to establish a methodology to test whether twitter sentiment is informative in predicting stock price movement and its volatility for the Chevron Corporation stock(CVX). This is done by using a DistilBert transformer architecture to assign a sentiment score for each tweet and afterwords, we try to determine if there is a correlation between stock price and daily sentiment using an ARMA GARCH Model.

\jy{alphabetical; no repeating words in title. Read my writing tips and practice them}
\bigskip
\noindent{\sc Keywords}:
BERT;
chevron;
twitter;
nlp;
sentiment analysis.

@@

\jy{what do I recommend for linewidth?}

\jy{Two figure labels are multiply defined. Read the log and fix the warnings}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec: intro}
%\paragraph{Overview of topic}

@@

\jy{the figure x-labels are cutoff}

\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig1.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022}\label{fig:fig1}
  \end{center}
\end{figure}

While the Highs and Lows might be informative in analysis, in order to concentrate the focus of this paper in a manner more similar to other financial data analyses, we will only experiment the Closing Price and the Returns which can be taken by taking the log difference of the Closing Prices.

\jy{The upper panel is repeating. Only keep one.}
\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig2.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022 and the returns}\label{fig:fig2}
  \end{center}
\end{figure}

\jy{Use table styles from my template}
We also used the data processing library Pandas in order to generate summary statistics as shown below:
\begin{table}[!hb]
  \begin{center}
    \caption{Summary Statistics for the Chevron Stock data}
    \label{tab:table1}
    \begin{tabular}{l|c|r} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Metrics} & \textbf{Close} & \textbf{Returns}\\
      \hline
      count & 503.000000 & 502.000000\\
      mean & 131.520815 & 0.001496\\
      std & 30.228298 & 0.018412\\
      min & 84.709999	 & - 0.069374\\
      25\% & 103.989998 & -0.008429\\
      50\% & 118.790001 & 0.002891\\
      75\% & 160.619995 & 0.013058\\
      max & 188.050003 & 0.085292\\
    \end{tabular}
  \end{center}
\end{table}

@@

\jy{reference to figures/tables by their labels. Never H or h. Always tbp for floating locations.}
\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig4.pdf}
  \caption{Garch Model Residuals}\label{fig:fig1}
  \end{center}
\end{figure}

@@

and those that remained were checked by the rugarch\citep{rugarch}, readr\citep{readr}, forecast\citep{forecast}, tseries\citep{tseries}, garchx\citep{garchx}, and TSA\citep{TSA} packages.
\jy{space before citep}

@@

As mentioned before, this paper is the first one to take the approach of using Transformers to determine whether twitter sentiment is informative in predicting the volatility of the stock price of the Chevron Corporation. In summary, this paper accomplished this task through the following:
\begin{itemize}
    \item Introducing some background behind the Chevron Corporation and Sentiment Analysis as a Practice.
    \item Describing the data and the methods by which the data was obtained.
    \item Describing the methods behind the sentiment analysis and correlation derivation.
    \item Describing the results.
\end{itemize}

\jy{close figure/table captions with a period.}

\jy{no need for paragraph titles}
\paragraph{Possible Expansions}
One possible expansion of this study would be to increase the sample size to include more observations and thus increase the robustness of the analysis. This would allow for more accurate statistical tests and a better understanding of the relationship between sentiment and stock price movement/volatility. Another possible expansion would be to test for causality using techniques like the Granger causality test, which would allow us to determine whether sentiment scores are truly predictive of stock price movements or whether they are just correlated. Additionally, conducting a meta-analysis of multiple companies in the fossil fuel industry or across multiple industries could provide broader insights into the relationship between sentiment and stock prices.

\paragraph{Takeaways}
Although this study found that both weighted and unweighted sentiment scores may not have been effective in predicting stock price movement and volatility of CVX, it established a novel methodology for analyzing Twitter sentiment data and determining whether it is informative in predicting stock prices. This methodology can be applied to other industries and sectors to gain a better understanding of the role that sentiment plays in stock price movements. The methodology can be useful for multiple groups, such as Public Relations departments in companies like Chevron, who can use this to assess the effectiveness of their publicity campaigns on the company's valuation, leading to higher profits and public influence. Alternatively, environmental groups can use this methodology to measure the effectiveness of their activism campaigns in influencing the trading power of Chevron Stock ownership. This methodology provides a way for various groups to adjust their strategies to influence public sentiment and in turn stock value.

\paragraph{Conclusion}
Overall, this study provides valuable insights into the use of Twitter sentiment analysis as a tool for predicting stock prices, and points towards potential avenues for further research. With further research, determined but analyzing the effect of public sentiment on stock prices of fossil fuel companies can provide valuable insight into how public opinion can be used to augment or reduce the effects of climate change.

>>>>

\begin{abstract}
As a publicly traded company, the Chevron Corporation has the 11th highest market capitalization in the world with 361.57 billion dollars in traded stocks. There exist many social factors that contribute to this stock value but this paper aims to establish a methodology to test whether public opinion in social media is informative in predicting stock price movement and its volatility for the Chevron Corporation stock(CVX). This is done by using a DistilBert transformer architecture to assign a sentiment score for each tweet and afterwords, we try to determine if there is a correlation between stock price and daily sentiment using an ARMA GARCH Model.

\jy{alphabetical; no repeating words in title. Read my writing tips and practice them}
%% did this
\bigskip
\noindent{\sc Keywords}:
BERT;
chevron;
twitter;
nlp;
sentiment analysis.

@@

\jy{what do I recommend for linewidth?}
%% need the ebook
\jy{Two figure labels are multiply defined. Read the log and fix the warnings}
%% could not find the figures
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec: intro}
%\paragraph{Overview of topic}

@@

\jy{the figure x-labels are cutoff}
%%% fixed this
\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig1.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022}\label{fig:fig1}
  \end{center}
\end{figure}

While the Highs and Lows might be informative in analysis, in order to concentrate the focus of this paper in a manner more similar to other financial data analyses, we will only experiment the Closing Price and the Returns which can be taken by taking the log difference of the Closing Prices. sample

\jy{The upper panel is repeating. Only keep one.}
%%%fixed this part
\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig2.pdf}
  \caption{A plot of historic CVX stock data from 2021 to 2022 and the returns}\label{fig:fig2}
  \end{center}
\end{figure}

\jy{Use table styles from my template}
%%% Need link to the efor,
We also used the data processing library Pandas in order to generate summary statistics as shown below:
\begin{table}[!hb]
  \begin{center}
    \caption{Summary Statistics for the Chevron Stock data}
    \label{tab:table1}
    \begin{tabular}{l|c|r} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      \textbf{Metrics} & \textbf{Close} & \textbf{Returns}\\
      \hline
      count & 503.000000 & 502.000000\\
      mean & 131.520815 & 0.001496\\
      std & 30.228298 & 0.018412\\
      min & 84.709999	 & - 0.069374\\
      25\% & 103.989998 & -0.008429\\
      50\% & 118.790001 & 0.002891\\
      75\% & 160.619995 & 0.013058\\
      max & 188.050003 & 0.085292\\
    \end{tabular}
  \end{center}
\end{table}

@@

\jy{reference to figures/tables by their labels. Never H or h. Always tbp for floating locations. }
%% Confused
\begin{figure}[!hb]
  \begin{center}
  \includegraphics[width=\textwidth]{../figures/fig4.pdf}
  \caption{Garch Model Residuals}\label{fig:fig1}
  \end{center}
\end{figure}

@@

and those that remained were checked by the rugarch \citep{rugarch}, readr \citep{readr}, forecast \citep{forecast}, tseries \citep{tseries}, garchx \citep{garchx}, and TSA \citep{TSA} packages.
\jy{space before citep}
%%did it

@@

As mentioned before, this paper is the first one to take the approach of using Transformers to determine whether twitter sentiment is informative in predicting the volatility of the stock price of the Chevron Corporation. In summary, this paper accomplished this task through the following:
\begin{itemize}
    \item Introducing some background behind the Chevron Corporation and Sentiment Analysis as a Practice.
    \item Describing the data and the methods by which the data was obtained.
    \item Describing the methods behind the sentiment analysis and correlation derivation.
    \item Describing the results.
\end{itemize}.

\jy{close figure/table captions with a period.}
%%think I did it but not sure

\jy{no need for paragraph titles}
%%done
\paragraph{}
One possible expansion of this study would be to increase the sample size to include more observations and thus increase the robustness of the analysis. This would allow for more accurate statistical tests and a better understanding of the relationship between sentiment and stock price movement/volatility. Another possible expansion would be to test for causality using techniques like the Granger causality test, which would allow us to determine whether sentiment scores are truly predictive of stock price movements or whether they are just correlated. Additionally, conducting a meta-analysis of multiple companies in the fossil fuel industry or across multiple industries could provide broader insights into the relationship between sentiment and stock prices.

\paragraph{}
Although this study found that both weighted and unweighted sentiment scores may not have been effective in predicting stock price movement and volatility of CVX, it established a novel methodology for analyzing Twitter sentiment data and determining whether it is informative in predicting stock prices. This methodology can be applied to other industries and sectors to gain a better understanding of the role that sentiment plays in stock price movements. The methodology can be useful for multiple groups, such as Public Relations departments in companies like Chevron, who can use this to assess the effectiveness of their publicity campaigns on the company's valuation, leading to higher profits and public influence. Alternatively, environmental groups can use this methodology to measure the effectiveness of their activism campaigns in influencing the trading power of Chevron Stock ownership. This methodology provides a way for various groups to adjust their strategies to influence public sentiment and in turn stock value.

\paragraph{}
Overall, this study provides valuable insights into the use of Twitter sentiment analysis as a tool for predicting stock prices, and points towards potential avenues for further research. With further research, determined but analyzing the effect of public sentiment on stock prices of fossil fuel companies can provide valuable insight into how public opinion can be used to augment or reduce the effects of climate change.
